---
title: "Arctic Data Center Dataset Citation Analysis"
author: "Althea Marks"
date: '2022-05-24'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(jsonlite)
library(tidyr)
library(dplyr)
library(dataone)
library(job)
```

## Run ADC DOIs through `scythe` & compare to known DataONE metrics citations

### Methods:

1) Gather existing/known ADC dataset citations picked up by the automated DataONE metrics API
2) Get a list of all ADC dataset DOIs
3) Run all ADC dataset DOIs through `scythe` libraries
4) Compare citations from `scythe` to DataONE metrics

#### 1) Retrieve all ADC citations with GET API request with the following request body

metrics service production endpoint: https://logproc-stage-ucsb-1.test.dataone.org/metrics


```{r get_request_citations, eval=FALSE}
{
  "metricsPage":{
    "total":0,
    "start":0,
    "count":0
  },
  "metrics":["citations"],
  "filterBy":[{
    "filterType":"repository",
    "values":["urn:node:ARCTIC"],
    "interpretAs":"list"
  },
  {
    "filterType":"month",
    "values":["01/01/2012",
              "05/24/2022"],
    "interpretAs":"range"
  }],
  "groupBy":["month"]
}
```

Example request:

```{r example_request, eval=FALSE}
https://logproc-stage-ucsb-1.test.dataone.org/metrics?metricsRequest={%22metricsPage%22:{%22total%22:0,%22start%22:0,%22count%22:0},%22metrics%22:[%22citations%22],%22filterBy%22:[{%22filterType%22:%22repository%22,%22values%22:[%22urn:node:ARCTIC%22],%22interpretAs%22:%22list%22},{%22filterType%22:%22month%22,%22values%22:[%2201/01/2012%22,%2205/24/2022%22],%22interpretAs%22:%22range%22}],%22groupBy%22:[%22month%22]
}
```

Function to retrieve dataset citations generated by DataOne metrics service (from Jeanette Clark):

```{r adc_metrics_citations_function}

metrics_citations <- function(from = as.POSIXct("2010-01-01"), to = as.POSIXct(Sys.Date())){

    from <- as.Date(from); to <- as.Date(to)
    from_q <- paste(stringr::str_pad(lubridate::month(from), 2, side = "left", pad = "0"),
                    stringr::str_pad(lubridate::day(from), 2, side = "left", pad = "0"),
                    stringr::str_pad(lubridate::year(from), 2, side = "left", pad = "0"),
                    sep = "/")

    to_q <- paste(stringr::str_pad(lubridate::month(to), 2, side = "left", pad = "0"),
                  stringr::str_pad(lubridate::day(to), 2, side = "left", pad = "0"),
                  stringr::str_pad(lubridate::year(to), 2, side = "left", pad = "0"),
                  sep = "/")

    d <- jsonlite::fromJSON(paste0('https://logproc-stage-ucsb-1.test.dataone.org/metrics?q={%22metricsPage%22:{%22total%22:0,%22start%22:0,%22count%22:0},%22metrics%22:[%22citations%22],%22filterBy%22:[{%22filterType%22:%22repository%22,%22values%22:[%22urn:node:ARCTIC%22],%22interpretAs%22:%22list%22},{%22filterType%22:%22month%22,%22values%22:[%22', from_q,'%22,%22', to_q, '%22],%22interpretAs%22:%22range%22}],%22groupBy%22:[%22month%22]}'))

    output_json <- d$resultDetails$citations # pulls citation info
    output_df <- as.data.frame(do.call(rbind, output_json), row.names = FALSE) # binds nested cit info into dataframe
   # output_clean <- rownames_to_column(output_df, var = "citation_id") # converts row names to column
    return(output_df)
}
```


```{r adc_citations}
# Run ADC API Get call, unnest target_id results to individual columns
cit_2022 <- metrics_citations()

cit_2022 <- tidyr::unnest(cit_2022,
                          cols = c(target_id, source_id, source_url,
                                   link_publication_date, origin, title,
                                   publisher, journal, volume, page, year_of_publishing))

```

#### 2) Get all ADC DOIs from SOLR query

DataOne metrics API can only provide data package DOIs with citations, and can not provide a comprehensive list of all data package DOIs contained within the ADC. To search through all the repository metadata we query the DataONE search index (Apache SOLR search engine). SOLR is the same underlying mechanism that DataONE uses in the online tool and can create complex logical query conditions.

```{r}
# complete list of searchable values
#getQueryEngineDescription(cn, "solr")

# set coordinationg node
cn <- dataone::CNode("PROD")

# point to specific member node
mn <- dataone::getMNode(cn, "urn:node:ARCTIC")

# set up Solr query parameters
queryParamList <- list(q="id:doi*", 
                       fl="id,title,dateUploaded,datasource",
                       start ="0",
                       rows = "100000")
# use `q = "identifier:doi* AND (*:* NOT obsoletedBy:*)"` to only include current versions of data packages 

# send query to Solr, return results as dataframe
result <- dataone::query(mn, solrQuery=queryParamList, as="data.frame", parse=FALSE)
```

#### 3) Run all ADC dataset DOIs through `scythe` libraries

```{r}
# create vector of all ADC DOIs from solr query `result`
adc_all_dois <- c(result$id[1:10])
length(adc_all_dois)

# install development version of scythe package with added xdd library
devtools::install_github("dataoneorg/scythe@develop")

```

- Search for ADC DOIs in `scythe`libraries

```{r}
# run each library search in parallel in separate background jobs
# by default job::job({}) imports everything into the job

job::job({
  adc_plos <- scythe::citation_search_plos(adc_all_dois)
})

job::job({
  adc_scopus <- scythe::citation_search_scopus(adc_all_dois)
  })

job::job({
  adc_springer <- scythe::citation_search_springer(adc_all_dois)
})

job::job({
  adc_xdd <- scythe::citation_search_xdd(adc_all_dois)
})

# Use this as backup for an overnight run
all <- scythe::citation_search(adc_all_dois, sources = c("plos", "scopus", "springer", "xdd"))
```

